
---
title: "Lab 5"
author: "Jaden Birkner"
date: "11/5/2019"
output: html_document
---

###Owner’s GitHub username: jebirkner
###Partner’s GitHub username: Mblount2
###Repo Name: ds202_lab5

```{r}
library(tidyverse)
```


1.
```{r}
diabetes = read.table('diabetes.txt',header=TRUE)
head(diabetes)
```

2.
```{r}
diabetes = diabetes %>% 
  mutate(frame = replace(frame, frame == '', NA))
```

```{r}
which(diabetes$frame == "")
```

```{r}
diabetes = droplevels(diabetes, exlude = levels(diabetes$frame==""))
```

```{r}
levels(diabetes$frame)
```

3.
```{r}
diabetes_reduced = diabetes %>% select(-c('id', 'bp.2s', 'bp.2d'))
head(diabetes_reduced)
```

4.

```{r}
diabetes_clean = diabetes_reduced %>% drop_na()
str(diabetes_clean)
```

5.
```{r}
sum(is.na(diabetes_clean)==TRUE)
```


6.
```{r}
ggplot(diabetes_clean, aes(x=glyhb)) + geom_histogram(binwidth = 1)
```

```{r}
ggplot(diabetes_clean, aes(x=log(glyhb))) + geom_histogram(binwidth = .25)
```

Potential Downside:

7.
```{r}
diabetes_clean$glyhb_star = log(diabetes_clean$glyhb)

ggplot(diabetes_clean, aes(x=glyhb_star)) + geom_histogram(binwidth = .25)
```

8.

```{r}
diabetes_clean$age_grouping <- cut(diabetes_clean$age, c(0,25,35,45,55,65,max(diabetes_clean$age)))
```

```{r}
diabetes_clean %>% group_by(age_grouping) %>% summarise(mean.glyhb = mean(glyhb_star))
```

It seems that there may be a positive correlation between age and glycosolated hemoglobin.

```{r}
diabetes_clean$weight_grouping <- cut(diabetes_clean$weight, c(5))
```

```{r}
diabetes_clean %>% group_by(weight_grouping) %>% summarise(mean.glyhb = mean(glyhb_star))
```

It seems that there may be a positive correlation between weight and glycosolated hemoglobin.

```{r}
diabetes_clean %>% group_by(gender) %>% summarise(mean.glyhb = mean(glyhb_star))
```

There does not seem to be a relationship between gender and glycosolated hemoglobin.

```{r}
diabetes_clean$waist_grouping <- cut(diabetes_clean$waist, c(5))
diabetes_clean %>% group_by(waist_grouping) %>% summarise(mean.glyhb = mean(glyhb_star))
```

It seems that there may be a positive correlation between waist size and glycosolated hemoglobin.

```{r}
diabetes_clean$chol_grouping <- cut(diabetes_clean$chol, c(10))
diabetes_clean %>% group_by(chol_grouping) %>% summarise(mean.glyhb = mean(glyhb_star))
```

It seems that there may be a positive correlation between cholesterol and glycosolated hemoglobin.

10.
```{r}
frame_mean = diabetes_clean %>% group_by(frame,location) %>% summarise (mean.glyhb_star= mean(glyhb_star))
frame_mean
```

```{r}
frame_mean %>% ggplot(aes(x = frame, y = mean.glyhb_star, color=location)) + geom_point()
```

